---
title: "6: Time Series Forecasting"
format: 
  html: 
    embed-resources: true
editor: visual
---

### Introduction

We're going to introduce time series forecasting and several packages that make forecasting easy.

A time series in econometrics refers to a sequence of data points that are collected and recorded over regular intervals of time, such as days, months, quarters, or years. These data points are typically organized in chronological order, and they represent the values of a specific variable or set of variables observed at different points in time.

Time series data is commonly used in economics and other fields to analyze and model various economic and social phenomena that exhibit temporal patterns and trends. Examples of variables that can be studied using time series analysis in econometrics include economic indicators (e.g., GDP, inflation rate, unemployment rate), financial market prices (e.g., stock prices, exchange rates), population statistics, and more. A simple example of temporal patterns is seasonality. For example, if you own a ski shop in Vail, Colorado, your sales will be higher during snowy months and lower during the summer. That's seasonality.

Time series analysis in econometrics involves various techniques and methods aimed at understanding the underlying patterns, trends, seasonal variations, cyclic behaviors, and other dynamics present in the data. These techniques include statistical methods like moving averages, exponential smoothing, autoregressive integrated moving average (ARIMA) models, and more advanced methods like vector autoregression (VAR), cointegration analysis, and state space models. The goal of time series analysis is often to make predictions about future values, identify potential relationships with other variables, and extract meaningful insights for decision-making and policy formulation.

Here, we'll do something fun to develop some intuition. We'll be "nowcasting" - analyzing Google trends data and forecasting where we think the trend will be in the immediate future. Why do that? Consider a scenario wherein you are selling a trending product, and you need to be able to forecast demand for your product for the next 3, 6, 9, 12 months. One way to do that is to use Google trends data to gauge how well your product will sell in the coming months, which in turn allows you to make sure you have the inventory, staff, and product on hand to maximize your profit efficiently.

### Install and load packages

There will be several new packages we'll be using today, so you'll need to install them per usual, e.g. `install.packages("timetk")` and so on.

```{r}
#| label: load-packages
#| warning: false
#| message: false

library(gtrendsR)
library(tidymodels)
library(modeltime)
library(tidyverse)
library(timetk)
library(lubridate)
```

### Google trends data

We can search for and download search term activity from Google Cloud. We're going to create a `search_term` to pass our query to Google, designate a geographical region from where we want the search to originate, and a time during which we're looking for that search term. We'll use the `gtrendsR` package to do this. Let's run a query for [Kia](https://www.kia.com/us/en), a carmaker who was known for cheap imports to the U.S. years ago, but has become a "cool brand" more recently.

Once we create the parameters, we can run the query, which will create an R object `gtrends_list` and contain monthly total "hits" for whatever we're searching for. The object also includes `interest_by_region`, which may be of more interest to physical shops, for example, and can be visualized separately.

```{r}
#| label: google-trends
#| warning: false
#| eval: false

# create search term, parameters for query
search_term <- "Kia"
location    <- "US"
time        <- "all"

# run Google trends query
gtrends_list <- gtrends(
    keyword      = search_term,
    geo          = location,
    time         = time,
    onlyInterest = TRUE
)
```

Click the swivel button in the Data pane to see what's in `gtrends_list`. As of this writing, we have over 240 monthly observations of 7 variables. I saved this file for reproducibility, but obviously if you're "nowcasting" you'll want to run fresh queries every time. I saved the trends data in a data object, an `.rds` file, which is a file format used to store R objects. If you want to reproduce my results below, use `gtrends_list.rds` instead of running the query yourself. The file is included in the folder for this lesson.

```{r}
#| label: save-rds
#| warning: false
#| eval: false

# Save for reproducibility
gtrends_list |>
    write_rds("gtrends_list.rds")
```

```{r}
#| label: load-rds
#| warning: false

# load saved file for this exercise
gtrends_list <- read_rds("gtrends_list.rds")
```

### A quick visualization

Let's plot the time series of "hits" per month for our search term. From our trends object, we're going to `pluck` the `interest_over_time` and visualize the number of hits. We'll also do a quick data cleaning step - drop zeros, impute missing values, and pull the cleaned series into a plotting-ready tibble.

```{r}
#| label: clean-and-plot
#| warning: false

gtrends_interest_over_time_tbl <- gtrends_list |>
    pluck("interest_over_time") |>
    as_tibble() |>
    select(date, hits) |>
    mutate(hits = ifelse(hits == 0, NA, hits),
           hits = ts_impute_vec(hits, period = 12))

# visualization
gtrends_interest_over_time_tbl |>
    plot_time_series(date, hits)

# cleaned data to a new tibble, standard naming
data_prepared_tbl <- gtrends_interest_over_time_tbl |>
    rename(value = hits)
```

Because there were few zeros if any in the "hits" field the visualization doesn't change much. That's okay. It's still important that you know to do a data cleaning step because other search terms are bound to have lots of zeros.

We now have a new cleaned data tibble `data_prepared_tbl` that we will use for modeling and forecasting.

### Modeling time series

What we want to do is forecast the future of the time series. To do that, we first need to split the data into a "training" set and a "test" set. We'll create a bunch of predictive models using the training data to calibrate the models, then "forecast" on the "test" set. Since we know the "test" data are actual observations, we can measure how the models perform in forecasting what has already happened. Once we have a sense of which models perform well, we can then use them on the whole data series to forecast the future with some degree of accuracy.

Let's create our splits into training and testing periods. We'll chop off the last year of data as the testing subset, and then take a look at what the training and test sets look like.

```{r}
#| label: split-series
#| warning: false

splits <- time_series_split(data_prepared_tbl, assess = "1 year", cumulative = TRUE)

# visualize the subsets
splits |>
    tk_time_series_cv_plan() |>
    plot_time_series_cv_plan(date, value)
```

We're going to use a number of models from the [`tidymodels`](https://www.tidymodels.org/) and [`modeltime`](https://business-science.github.io/modeltime/) packages. Each model will be expressly called and trained on the training subset of our time series data. The coding varies only slightly, and there are many options to fine tune your models. Refer to the documentation for more information. For now, we're just going to see what these things do, with very brief explanations of the models themselves. In each case, we will create a "fit" object wherein we fit the model to the data. We will then evaluate the fit by testing it on the test subset of our data, the last year of the time series. We tell R which model to fit by calling a particular engine with `set_engine` and tell it to run the model on the training data with `data = training(splits)`. Finally, in each model there will be a formula that the model will fit in the form of `dependent variable ~ independent variables`. On the left hand side of the `~` is the dependent variable you are trying to explain and on the right hand side you have the variables you are using to try to explain the dependent variable.

#### ARIMA

ARIMA stands for Auto-Regressive Integrated Moving Average. It's a widely used time series forecasting model in econometrics and statistics that combines auto-regressive (AR) and moving average (MA) components with differencing to handle non-stationary data. ARIMA models are particularly useful for analyzing and predicting time series data that exhibit trends, seasonality, and autocorrelation. We know what trends and seasonality are, but in case you're unfamiliar with autocorrelation, it is the tendency for past data to predict future data. If we have 47 days of extreme heat, chances are tomorrow will be hot, too. Similarly, if January is cold and July is hot, if the date is a December date we can predict that a future July date will be hot. That, poorly explained, is autocorrelation. Note that the only "explainer" variable in the ARIMA model is the date itself.

```{r}
#| label: arima-fit
#| warning: false

model_fit_arima <- arima_reg() |>
    set_engine("auto_arima") |>
    fit(value ~ date, data = training(splits))
```

#### Linear models

Linear models calculate the correlation between a dependent variable - in our case, the number of hits in a month - and one or more independent, or "explanatory" variables. For example, if we had lung cancer data, we might model the incidence of cancer, the dependent variable, on a number of independent variables, like age, sex, race, physical fitness, and tobacco use. That would be a "multiple linear regression" model, because we have multiple independent variables that may increase or decrease the likelihood that someone develops cancer. In our case, we only have previous observations of the number of hits, as well as the months, as independent variables. We'll see if linear models are any good at predicting search data here. We'll use two here, one with and one without a trend variable.

```{r}
#| label: linear-models
#| warning: false

# linear regression with trend
model_fit_lm <- linear_reg() |>
    set_engine("lm") |>
    fit(value ~ as.numeric(date) + month(date, label = TRUE),
        data = training(splits))

# linear regression without a trend variable
model_fit_lm_no_trend <- linear_reg() |>
    set_engine("lm") |>
    fit(value ~ month(date, label = TRUE),
        data = training(splits))
```

#### Prophet

Prophet is a forecasting model developed by Facebook that is specifically designed for time series forecasting with an emphasis on handling various components of time series data such as trends, seasonality, holidays, and special events. It is particularly useful for forecasting time series data that exhibit complex patterns and multiple sources of variation.

Prophet can handle missing data, outliers, and abrupt changes in the time series. The model's flexibility allows it to adapt to various types of time series data without extensive pre-processing. You can also fine-tune your model with the inclusion of other variables that might contribute to the data you are observing.

```{r}
#| label: prophet
#| warning: false

model_fit_prophet <- prophet_reg() |>
    set_engine("prophet") |>
    fit(value ~ date, data = training(splits))
```

#### XGBOOST

XGBOOST is, for lack of a better term, a supermodel. Short for eXtreme Gradient Boosting, the model is known for very quick computation on large data sets, its ability to handle complex relationships in data, and accurate predictions in both classification and regression situations. It wins a lot of [Kaggle](https://www.kaggle.com/) competitions.

The algorithm starts with a simple tree model, then looks at the errors of its predictions compared to the observed variables, then reruns itself to try to predict the residuals, etc. It's pretty smart.

```{r}
#| label: xgboost
#| warning: false

model_fit_xgboost <- boost_tree(mode = "regression") |>
    set_engine("xgboost") |>
    fit(value ~ as.numeric(date) + month(date, label = TRUE),
        data = training(splits))
```

There are many more time series forecasting models available to data scientists, but you get the idea. Once you learn more about these and your other options, you will be in a better position to meaningfully implement forecasting models with your own specific data and needs. What you do *not* want to do, however, is throw a bunch of models at your data without understanding if they are appropriate, theoretically, for your particular situation. Not every model is appropriate for every forecasting scenario, just as surgery is not the answer for every malady. For now, let's see how these models perform on our Kia data.

### Testing...testing...

Now that we have our fitted models, it's time to see how they do in "predicting" the last year of our actual data (the testing subset from `splits`). First, let's create a table to pass to our calibration process.

```{r}
#| label: modeltime-table

model_tbl <- modeltime_table(
    model_fit_arima,
    model_fit_lm,
    model_fit_lm_no_trend,
    model_fit_prophet,
    model_fit_xgboost
)
```

Now we're going to calibrate our models such that they "predict" the last year of data, passing the models from the table we just created to a new table of results. We'll then visualize what the "forecasts" look like compared to the actual data in that last year.

```{r}
#| label: calibration-table
#| warning: false

calibration_tbl <- model_tbl |>
    modeltime_calibrate(testing(splits))

calibration_tbl |>
    modeltime_accuracy() |>
    table_modeltime_accuracy(resizable = TRUE, bordered = TRUE)

calibration_tbl |>
    modeltime_forecast(
        new_data      = testing(splits),
        actual_data   = data_prepared_tbl,
        conf_interval = 0.80
    ) |>
    plot_modeltime_forecast(.legend_show = TRUE, .legend_max_width = 25)
```

Open the plotly in a new window. Zoom in to the forecast area and compare it to the actual data. It looks like XGBOOST and Prophet often do the best overall, with the smallest errors (mae, mape, rmse, etc.) in the calibration table of results. Now we're going to refit the models to the entire dataset, then forecast forward one year. Again we'll use a confidence interval of 80%. We will create a new plot with the forecasts into the future.

```{r}
#| label: refit-forecast
#| warning: false

# refit
refit_tbl <- calibration_tbl |>
    modeltime_refit(data = data_prepared_tbl)

# forecast table
forecast_tbl <- refit_tbl |>
    modeltime_forecast(
        h             = "1 year",
        actual_data   = data_prepared_tbl,
        conf_interval = 0.80
    )

# create plotly
forecast_tbl |>
    plot_modeltime_forecast(.interactive = TRUE)
```

You can again open and expand the window, zoom in on the forecasts, etc. For your actual project, given we see that the linear model with no trend is performing poorly, you'd likely go back several steps, drop that model, perhaps run another model, and assess whether to keep some or all of your models.

### Model averaging

You want a group of well-performing models because we're now going to create an average of the model forecasts into a "super" forecast, so we're working with just one forecast series. The average is supposed to be better than any one model, and we want the most accurate forecast from which to plan real operations, e.g. ordering inventory, adjusting staff schedules, hiring, messaging, etc.

```{r}
#| label: model-averaging
#| warning: false

mean_forecast_tbl <- forecast_tbl |>
    filter(.key != "actual") |>
    summarise(across(.value:.conf_hi, mean),
              .by = c(.key, .index)) |>
    mutate(.model_id   = 99,
           .model_desc = "average of models")

# visualize mean forecast
forecast_tbl |>
    filter(.key == "actual") |>
    bind_rows(mean_forecast_tbl) |>
    plot_modeltime_forecast()
```

Since the mean forecast includes the linear model with no trend, which was well below the other model forecasts, the aggregated forecast is "dragged down" somewhat because of the inclusion of the lm no trend model. As noted above, if you were doing this for your company you would likely go back and drop (and perhaps replace) that model in lieu of something that seemed to work better. For now, we've done a lot here. Well done.

### Looking ahead

In the next lesson we move from forecasting a single series into full-blown tidy modeling on larger, multivariate datasets - and build a little app to explain a model's predictions to a non-technical audience.
